---
title: "R and Application"
author: "Arjun Dutta"
date: "13 August 2018"
output: html_document
---
# A.R Programming

####1. R Input and Evaluation

"x <- 1" This is an expression

"x" Autoprinting

"print(x)" explicit printing

"#" for comment

":" used to create integer sequence

"str(object)" is diagonostic function is used to compactly display the internal structure of an R object.

"split(airquality,airquality$Month)" to split that dataset by month into list which contains 5 dataframes.

"gl(n,k,length=n*k,labels=seq_len(n),ordered=F)" generates factors by specifying the patterns of their level

n is integer giving the number of levels 
k is integer giving the number of replication

#####Vectorised Operation

Operations in R is vectorised

```{r , include =FALSE}
# All are element wise
x <- 1:4
y <- 5:8
x + y
x - y
x * y
x / y

matrix <- matrix(1:4,2,2)
matriy <- matrix(5:8,2,2)
matrix * matriy #element wise matrix multiplication
matrix/matriy   #element wise division
matrix %*% matriy  #true matrix multiplication

```

####2. Data Types: Objects and Attributes

R has 5 basic atomic classes of object:

* character
* numeric
* integer
* complex
* logical(True/False)

####3. Numbers

Numbers in R is treated as numeric objects(i.e. double precision real number)

specify "L" after an number to explicitly give you an integer

"1/0" is Inf

"0/0" is NaN or missing value

####4. Attributes

R objects are built upon attributes:

* names, dimnames
* dimension       Every matrices has row and col or multidimensional array
* class           Every object is a class
* length
* mode
* other user defined attributes/metadata 

"attributes()" function allows you to set or modify the attribute of an object.

####5. Data Types: Vector and List

* A vector is a basic objct that contain objects of same class
* list is represented a vector but can contain objects of different class

Vectors can be created using "c()" stands for "concatenate" and "vector(class="numeric",lenght=10)" functions will create a vector of with "0" values of length 10.

####6. Coercion

Implicit Coercion

Explicit Coercion

as.* function   i.e. as.integer(), as.character() etc

List are special type of vector that can contain elements of different classes. 

"x <- list(1,"a", TRUE, 1+4i)"

The elements are indexed by double bracket.

####7. Data Types: Matrices

Matrices are vector with dime=ension attribute. The Dimension attribute is itself an integer vector of length 2.

"m <- matrix(nrow = 2, ncol = 3)" will create a NA matrix

Matrices number are inserted coloumn wise unless byrows = T argument is specified

* Matrice by vector

"m <- 1:10
 dim(m) <- c(2,3)"

* Joining mutiple vector as Coloumn

"cbind(x,y)"

* Joining mutiple vector as Row

"rbind(x,y)"

str(matrix) returns two dimensional homogeneous array structure.

####8. Data Types: Factors

Factors are used to represent categorical data. Factors can be ordered and unordered. factors are integer vector where integers are labels.

"lm()" and "glm()" treats factors specially.

####9. Data Type: Missing Value

NA and NaN are Missing Values 

"is.na()" and "is.nan()" is used to test objects if they are NA and NaN and it returns a logical vector.

NA values have class also integers, character etc

NaN is NA but NA is not NaN

####10. Data Types: Dataframes

Data Frame is used to store tabular data is a special type of list has to have the same length

####11. Data Types: Names Attribute

R objects can also have names, which is very useful for writing readable codes and self describing objects.

* for vector
"x <- 1:3
names(x) <- c("A", "B", "C")"

* for list
"x <- list(a=1,b=2,c=3)"

* for matrices

"x <- matrices(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("A", "B"), c("C", "D"))"

####12. Reading Tabular Data

####Reading Data

* "read.csv(file= "File path/Connection", header = T/F, sep = ",", colClasses = , nrow) and read.table()" is used to read tabular data

* "readLines" for reading lines of a text file

* "source()" for reading R codes

* "dget()" for reading R code files but for reading R objects that has been dparsed into text file.

* "load() and unserialize()" is used to read R binary object.

#####Reading Large Data

* Know and specify the colClasses of the dataset everytime
This makes read.table run a lot faster.
This is quick and a dirty way to know the class fo any dataset
"
initial <- read.table("datatable.txt", nrows=10)
classes <- sapply(initial,class)
Data <- read.table("datatable.txt", colClasses = classes)
"

Also read.table(file,colClasses = "numeric") assumes all coloumns are numeric

####Writing Data

* "write.table()" "writeLines()" "dump" "dput" "save" "serialize" are paired up with their reading analog.

####Subsetting

Single bracket "[]" 
1. Always returns an object of same class as the original 
2. Can be used to select more than one variable.

Double Bracket"[[]]"
1. Used to extract element of a list or a data frame
2. It can be only used to extract a single element and the class of the returned object will not necessarily be a list or data frame.

Dollar sign "$"
1. Used to extract element of a list or a dataframe by name.

#####Indexing

Numeric indexing is used to extract element by number

Logical indexing is used to extract element that meet certain condition

#####Subsetting-List

```{r , include=FALSE}
x <- list(foo=1:4, bar = 0.6, baz = "Hello")
x[1]   #Returns List
x[[1]] #Returns integer vector sequence
x$bar  #Returns integer vector sequence
x["bar"] #Returns List
x[["bar"]] #Returns numeric since only one value

x$foo[1] #extracting single element
x[c(1,3)] #use single bracket to extract multiple element of a list

name <- "foo"
x[[name]]     #"[[]]" can be used with computed names

x[[c(1,3)]] #double bracket is used to extract nested elements

```

#####Subsetting-Matrices

```{r , include=FALSE}

x <- matrix(1:9,3,3)

x[1,2] #Extract elements integer

x[,1] #1st col integer
x[1,] #1st row integer

x[1:2,1:2] #Matrix class is returned

x[1,2, drop = FALSE] #matrix Class is returned
```

#####Subsetting-Partial Matching

```{r setup, include=FALSE}
x <- list(aardvark = 1:10)
x[["a"]]
x[["a", exact=F]] #Matching Variable 
```

#####Subsetting-Missing Values

```{r , include=FALSE}
x <- c(1,2,NA,5,6,NA)
bad <- is.na(x)   #is a logical vector true where there is a NA
x[!bad] #not operator for non missing values

#Removing NA from multiple vectors

x <- c(1,2,NA,5,6,NA)
y <- c(NA,NA,4,7,5,NA)
good <- complete.cases(x,y)
x[good]
y[good]

#Removing NA from Dataframe
good <- complete.cases(airquality)
airquality[good,][1:6,]
```

####Control Structure

IF Structure 

"
if(condition)
{
expression
}

else if(condition)
{
expression
}

else
{
expression
}
"

"
if(x>3)
{
y <- 10
}
else
{
y <- 0
}
"
is same as
"
y <- if(x>3)
{
  10
}
else
{
0
}
"

Sometime only the if clause works else clause might not be neceessary.

FOR Structure

####Function

Functions are "first class objects" which means that they can be treated much like any other R objects.

* Function can be pased as an argument to another function

* Functions can be nested i.e you can define a function inside another function

#####create a function 

"
obj_name <- function(arg1,arg2){
      expression
}

"

* You don't have to specify return because the last expression is always returned

* Objects in the argument of function can be any Data Structure i.e vector, list, dataframe, etc

Example

1. Create a function that extract number above a threshold from a vector

"
above_n <- function(x, n=10){
      use <- x > n
      x[use]
}

x <- 1:20

above_n(x, 15)

"

n=10 in the function is the default value of n

2. Create a function to calculate mean of coloumns of a data frame: -

"
col_mean <- function(y, removena = TRUE){
      nc <- ncol(y)
      means <- numeric(nc)
      for(i in 1:nc){
      means[i] <- mean(y[,i], na.rm = removena)
      }
      means
}

col_mean(airquality)

"

* "formals(takes just the name of the function)" return the arguments of the named function

#####



####Apply Family for Loop





####Simulation

Probability distribution functions have four function associated with them:

* d for density

* r for random number generation

* p for cumulative distribution i.e. P[x<=p]

* q for quantile distribution

dnorm(x, mean = 0, sd = 1, log = FALSE)
pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p  = FALSE)
rnorm(n, mean = 0, sd = 1)
 
If phi be the cumulative distribution function  =for standard normal distribution then pnorm(q) = phi(q) and qnorm(p) = phi-inverse(p)

when doing simulation always set the seed using 
set.seed(any number) if you want to generate the same set of random numbers again
because if you want to reproduce the same result that is why you use the set seed.

#####Simulating a Linear Model

* Suppose we want to simulate from the following Linear Model

Y = /* B^0 + B^1*X + */

where 

#####Random Sample

sample(1:10,4) will give 4 number from1 to 10
sample(letters,5) will give 5 small letters from a to z 
sample(LETTERS) will give 5 big letters from A to Z 
sample(1:10) will give 10 numbers from 1 to 10 in permutation
sample(1:10, replace = T) will give sample with replacement

####R Profiler

The Idea of Profiling is a systematic way to examine how much time is spend in different parts of program or you want to optimize your code when you are looping 1000 iterations.

You should first think about how to run the code how to amke it readable how other people can understand it then think of optimising your code.

system.time() to check the time returns an abject of class proc_time.

There are packages called Multi-Threaded BLAS(Basic Linear Algebra Standard) optimizes to use multiple cores

MKL for Intel Machines
Atlas for AMD

# B.Getting and Cleaning Data

####Downloading Files:

* Know the working directory

"getwd()" gets the working directory and "setwd()" sets the working directory

In windows you have to use double backslahes (\\) rather than forward slashes 

"file.exists("directoryname")" to check directory exist or not

"dir.create("directoryname")"" to create a directory

if(!file.exist("data")){
dir.create("data")
}

"download.file(url,destfile,method)" used to download csv,tab delimited 
if you are working in windows you don't need to specify method argument

# C.Exploratory Data Analysis



####